---
title: 'Temporal Variability in Pediatric population: SITEID'
output:
  html_document:
    code_folding: show
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---

```{r libraries, message=FALSE, warning=FALSE, results='hide'}
# Install packages
paket <- function(pak){
  new_pak <- pak[!(pak %in% rownames(installed.packages()))]
  if (length(new_pak)) 
    install.packages(new_pak, dependencies = TRUE,repos="https://cloud.r-project.org/",lib=userLibDir)
  sapply(pak, library, character.only = TRUE)
}

listOfPackages <- c("EHRtemporalVariability", "dplyr", "tidyr", "tidyverse", "tableone", "plotly", "DT")
paket(listOfPackages)
```

```{r}
theme_set(theme_bw() + 
            theme(legend.title = element_blank(),
                  panel.grid.minor = element_blank()))
```


## Data read-in

Read the functions and data needed for the analysis. 
```{r suppData, message=FALSE}
source("R/readInputFiles.R")
source("R/summaryTableOne.R")
autoimmune <- read.delim('public-data/autoimmuneICD10codes.txt')
mentalDisorder <- read.delim('public-data/icd10_categories_mentalDisorders.txt')
```

Read the input files.

```{r message=FALSE, warning=FALSE}
### Read the CSV input files
#change to the path where your phase 2.1 data files are located
myinputFiles <- "/4ceData/Input/"
files <- readInputFiles( path      = myinputFiles, 
                         separator = ",",
                         skip      = 0, 
                         verbose   = FALSE )
  
### Extract the patient summary and observation information. 
demo_raw <- files[[1]]
obs_raw <- files[[2]]
clinical_raw <- files[[3]]

### Determine the obfuscation threshold (FALSE if no obfuscation, the numeric value of the obfuscation threshold if any)
obfuscation =  FALSE
#obfuscation = 3

### Determine if you want to visualize counts or frequencies
absoluteValue = TRUE
```

# Mental Disorders
ICD10 codes extracted from: The ICD-10 Classification of Mental and Behavioural Disorders Clinical descriptions and diagnostic guidelines by the World Health Organization https://www.who.int/classifications/icd/en/bluebook.pdf

Suidice related codes added from: https://www.cdc.gov/nchs/data/nhsr/nhsr108.pdf 

```{r}
DT::datatable( mentalDisorder, rownames = FALSE )
```

```{r  message=FALSE, warning=FALSE}
observations <- obs_raw %>%
          filter( concept_type == 'DIAG-ICD10', days_since_admission >= 0, concept_code %in% mentalDisorder$ICD10) %>%
          mutate( pair = as.character(paste0( patient_num, "*",days_since_admission))) %>%
          select( patient_num, pair, concept_code )
```

```{r}
observations$concept_code <- as.character( observations$concept_code)
  
clinical <- clinical_raw %>%
    mutate( pair = as.character(paste0( patient_num, "*",days_since_admission))) %>%
    select( patient_num, pair, calendar_date, severe, deceased)
  
final <- merge( observations, clinical, by="pair") 
  
final <- final %>%
    mutate(patient_num = patient_num.x) %>%
    select(patient_num, concept_code, calendar_date, severe )
```

## Demographic summary

```{r}
## Select of demographic data
demographics <- unique( demo_raw %>%
                        select( patient_num, age_group, sex ) )

demographics$subset <- ifelse( demographics$patient_num %in% final$patient_num, "Subset Yes", "Subset No")
  
## Create the summary table 1
tbl1 <- summaryTableOne( df = demographics, obfuscation = obfuscation, stratifyBy = "Mental Disorders")
```

### Format the data for the temporal variability analysis
```{r}
## Remove the patient identifier and remove duplicates
final <- final[, -1 ]
final <- final[!duplicated( final ), ]
  
## Put the date column as class Date
final <- EHRtemporalVariability::formatDate(
    input         = final,
    dateColumn    = "calendar_date",
    dateFormat = "%Y-%m-%d"
)
```

### Add the description of the ICD codes for mental disorders
```{r}
colnames( mentalDisorder )[2] <- "description"
      final <- merge( final, mentalDisorder, by.x = "concept_code", by.y = "ICD10")
final$description <- as.character(final$description)
final$descriptionShort <- substr(final$description, 1, 40)
```



## Heatmap output visualization

### By week
```{r, warning=FALSE}
probMaps_week <- estimateDataTemporalMap(data         = final, 
                                    dateColumnName = "calendar_date", 
                                    period         = "week")

if( obfuscation !=  FALSE ){
    
    for( i in 1:length(probMaps_week)){
      probMaps_week[[i]]@countsMap[probMaps_week[[i]]@countsMap > 0 & probMaps_week[[i]]@countsMap <= obfuscation ] <- 0.5 * obfuscation
    }
}

plotDataTemporalMap(
    dataTemporalMap =  probMaps_week[["descriptionShort"]],
    startValue = 1,
    endValue = 20,
    colorPalette    = "Magma", 
    absolute = absoluteValue)

```


### By month
```{r, warning=FALSE}
probMaps_month <- estimateDataTemporalMap(data         = final, 
                                    dateColumnName = "calendar_date", 
                                    period         = "month")
if( obfuscation !=  FALSE ){
    for( i in 1:length(probMaps_month)){
      probMaps_month[[i]]@countsMap[probMaps_month[[i]]@countsMap > 0 & probMaps_month[[i]]@countsMap <= obfuscation ] <- 0.5 * obfuscation
    }
} 

plotDataTemporalMap(
    dataTemporalMap =  probMaps_month[["descriptionShort"]],
    startValue = 1,
    endValue = 20,
    colorPalette    = "Magma", 
    absolute = absoluteValue)
```

# Autoimmune disorders

```{r}
DT::datatable( autoimmune, rownames = FALSE )
```


```{r  message=FALSE, warning=FALSE}
obfuscation = FALSE
observations <- obs_raw %>%
        filter( concept_type == 'DIAG-ICD10', days_since_admission >= 0, concept_code %in% autoimmune$ICD10) %>%
        mutate( pair = as.character(paste0( patient_num, "*",days_since_admission))) %>%
        select( patient_num, pair, concept_code )
```

```{r}
observations$concept_code <- as.character( observations$concept_code)
  
clinical <- clinical_raw %>%
    mutate( pair = as.character(paste0( patient_num, "*",days_since_admission))) %>%
    select( patient_num, pair, calendar_date, severe, deceased)
  
final <- merge( observations, clinical, by="pair") 
  
final <- final %>%
    mutate(patient_num = patient_num.x) %>%
    select(patient_num, concept_code, calendar_date, severe )
```

## Demographic summary

```{r}
## Select of demographic data
demographics <- unique( demo_raw %>%
                        select( patient_num, age_group, sex ) )

demographics$subset <- ifelse( demographics$patient_num %in% final$patient_num, "Subset Yes", "Subset No")
  
## Create the summary table 1
tbl2 <- summaryTableOne( df = demographics, obfuscation = obfuscation, stratifyBy = "Autoimmune Disorders")
```

## Format the data for the temporal variability analysis
```{r}
## Remove the patient identifier and remove duplicates
final <- final[, -1 ]
final <- final[!duplicated( final ), ]
  
## Put the date column as class Date
final <- EHRtemporalVariability::formatDate(
    input         = final,
    dateColumn    = "calendar_date",
    dateFormat = "%Y-%m-%d"
)
```

### Add the description of the ICD codes for mental disorders
```{r}
colnames( autoimmune )[2] <- "description"
final <- merge( final, autoimmune, by.x = "concept_code", by.y = "ICD10")
final$description <- as.character(final$description)
final$descriptionShort <- substr(final$description, 1, 40)
```

## Heatmap output visualization

### By week
```{r, warning=FALSE}
probMaps_week <- estimateDataTemporalMap(data         = final, 
                                    dateColumnName = "calendar_date", 
                                    period         = "week")
if( obfuscation !=  FALSE ){
    for( i in 1:length(probMaps_week)){
      probMaps_week[[i]]@countsMap[probMaps_week[[i]]@countsMap > 0 & probMaps_week[[i]]@countsMap <= obfuscation ] <- 0.5 * obfuscation
    }
}

plotDataTemporalMap(
    dataTemporalMap =  probMaps_week[["descriptionShort"]],
    startValue = 1,
    endValue = 20,
    colorPalette    = "Magma", 
    absolute = absoluteValue)
```

### By month
```{r, warning=FALSE}
probMaps_month <- estimateDataTemporalMap(data         = final, 
                                    dateColumnName = "calendar_date", 
                                    period         = "month")
if( obfuscation !=  FALSE ){
    for( i in 1:length(probMaps_month)){
      probMaps_month[[i]]@countsMap[probMaps_month[[i]]@countsMap > 0 & probMaps_month[[i]]@countsMap <= obfuscation ] <- 0.5 * obfuscation
    }
}

plotDataTemporalMap(
    dataTemporalMap =  probMaps_month[["descriptionShort"]],
    startValue = 1,
    endValue = 20,
    colorPalette    = "Magma", 
    absolute = absoluteValue)
```

# Aggregated by PheWAS category
PheCode mapping file downloaded from: https://phewascatalog.org/phecodes_icd10 


```{r  message=FALSE, warning=FALSE}
observations <- obs_raw %>%
            filter( concept_type == 'DIAG-ICD10', days_since_admission >= 0) %>%
            mutate( pair = as.character(paste0( patient_num, "*",days_since_admission))) %>%
            select( patient_num, pair, concept_code )
```

```{r}
observations$concept_code <- as.character( observations$concept_code)
  
clinical <- clinical_raw %>%
    mutate( pair = as.character(paste0( patient_num, "*",days_since_admission))) %>%
    select( patient_num, pair, calendar_date, severe, deceased)
  
final <- merge( observations, clinical, by="pair") 
  
final <- final %>%
    mutate(patient_num = patient_num.x) %>%
    select(patient_num, concept_code, calendar_date, severe )
```

## Demographic summary

```{r}
## Select of demographic data
demographics <- unique( demo_raw %>%
                        select( patient_num, age_group, sex ) )

demographics$subset <- ifelse( demographics$patient_num %in% final$patient_num, "Subset Yes", "Subset No")
  
## Create the summary table 1
tbl3 <- summaryTableOne( df = demographics, obfuscation = obfuscation, stratifyBy = "PheWAS code")
```

## Format the data for the temporal variability analysis
```{r}
## Remove the patient identifier and remove duplicates
final <- final[, -1 ]
final <- final[!duplicated( final ), ]
  
## Put the date column as class Date
final <- EHRtemporalVariability::formatDate(
    input         = final,
    dateColumn    = "calendar_date",
    dateFormat = "%Y-%m-%d"
)
```

### Add the description of the ICD codes for mental disorders
```{r}
phecode <- read.csv("public-data/phecode_icd10.csv", colClasses = "character")

phecode$Phenotype <- ifelse( phecode$Phenotype == "", phecode$ICD10.String, phecode$Phenotype )
      phecode <- unique( phecode %>%
        mutate( concept_code = ICD10, description = Phenotype ) %>%
        select( concept_code, PheCode, description ))
      final <- merge( final, phecode)

final$description <- as.character(final$description)
final$descriptionShort <- substr(final$description, 1, 40)
```

## Heatmap output visualization

### By week
```{r, warning=FALSE}
probMaps_week <- estimateDataTemporalMap(data         = final, 
                                    dateColumnName = "calendar_date", 
                                    period         = "week")
if( obfuscation !=  FALSE ){
    
    for( i in 1:length(probMaps_week)){
      probMaps_week[[i]]@countsMap[probMaps_week[[i]]@countsMap > 0 & probMaps_week[[i]]@countsMap <= obfuscation ] <- 0.5 * obfuscation
    }
}

plotDataTemporalMap(
    dataTemporalMap =  probMaps_week[["descriptionShort"]],
    startValue = 1,
    endValue = 20,
    colorPalette    = "Magma", 
    absolute = absoluteValue)

save( probMaps_week, file="/4ceData/temporalVariability/export.RData")
```

### By month
```{r, warning=FALSE}
probMaps_month <- estimateDataTemporalMap(data         = final, 
                                    dateColumnName = "calendar_date", 
                                    period         = "month")
if( obfuscation !=  FALSE ){
    for( i in 1:length(probMaps_month)){
      probMaps_month[[i]]@countsMap[probMaps_month[[i]]@countsMap > 0 & probMaps_month[[i]]@countsMap <= obfuscation ] <- 0.5 * obfuscation
    }
}
plotDataTemporalMap(
    dataTemporalMap =  probMaps_month[["description"]],
    startValue = 1,
    endValue = 20,
    colorPalette    = "Magma", 
    absolute = absoluteValue)
```


# Medication

```{r  message=FALSE, warning=FALSE}
observations <- obs_raw %>%
      filter( concept_type == 'MED-CLASS', days_since_admission >= 0) %>%
      mutate( pair = as.character(paste0( patient_num, "*",days_since_admission))) %>%
      select( patient_num, pair, concept_code )
    observations$concept_code <- as.character(observations$concept_code)
```

```{r}
observations$concept_code <- as.character( observations$concept_code)
  
  clinical <- clinical_raw %>%
    mutate( pair = as.character(paste0( patient_num, "*",days_since_admission))) %>%
    select( patient_num, pair, calendar_date, severe, deceased)
  
  final <- merge( observations, clinical, by="pair") 
  
  final <- final %>%
    mutate(patient_num = patient_num.x) %>%
    select(patient_num, concept_code, calendar_date, severe )
```

## Demographic summary

```{r}
## Select of demographic data
demographics <- unique( demo_raw %>%
                        select( patient_num, age_group, sex ) )

demographics$subset <- ifelse( demographics$patient_num %in% final$patient_num, "Subset Yes", "Subset No")
  
## Create the summary table 1
tbl4 <- summaryTableOne( df = demographics, obfuscation = obfuscation, stratifyBy = "Medication")
```

## Format the data for the temporal variability analysis
```{r}
## Remove the patient identifier and remove duplicates
final <- final[, -1 ]
final <- final[!duplicated( final ), ]
  
## Put the date column as class Date
final <- EHRtemporalVariability::formatDate(
    input         = final,
    dateColumn    = "calendar_date",
    dateFormat = "%Y-%m-%d"
)
```

## Heatmap output visualization

### By week
```{r, warning=FALSE}
probMaps_week <- estimateDataTemporalMap(data         = final, 
                                    dateColumnName = "calendar_date", 
                                    period         = "week")

if( obfuscation !=  FALSE ){
    
    for( i in 1:length(probMaps_week)){
      probMaps_week[[i]]@countsMap[probMaps_week[[i]]@countsMap > 0 & probMaps_week[[i]]@countsMap <= obfuscation ] <- 0.5 * obfuscation
    }
}

plotDataTemporalMap(
    dataTemporalMap =  probMaps_week[["concept_code"]],
    startValue = 1,
    endValue = 20,
    colorPalette    = "Magma", 
    absolute = absoluteValue)
```

### By month
```{r, warning=FALSE}
probMaps_month <- estimateDataTemporalMap(data         = final, 
                                    dateColumnName = "calendar_date", 
                                    period         = "month")
if( obfuscation !=  FALSE ){
    for( i in 1:length(probMaps_month)){
      probMaps_month[[i]]@countsMap[probMaps_month[[i]]@countsMap > 0 & probMaps_month[[i]]@countsMap <= obfuscation ] <- 0.5 * obfuscation
    }
}

plotDataTemporalMap(
    dataTemporalMap =  probMaps_month[["concept_code"]],
    startValue = 1,
    endValue = 20,
    colorPalette    = "Magma", 
    absolute = absoluteValue)
```

# R session information
```{r}
sessionInfo()
```

